---
title: "IMDB Load name.basics.tsv.gz Data Set"
output: html_notebook
---


```{r}

#set this to the folder in which the file is located
#path <- "c:/Users/Martin/Downloads/imdb"
#path <- "c:/Users/mjs01/Downloads/imdb"
path <- "/Users/mschedlb/Downloads/imdb"

```


## Load a sample from 'name.basics.tsv'

This file contains actors, actresses, and others with their names and their roles
plus titles (movies, shows) they are "known for". 

Roles and knownFor are multi-valued string fields. Builds a second file
that maps one name to one knownFor title (nm,tt) to facilitate joins later.

The algorithm is to load the entire file:

* Updates the file and eliminates \\N values to empty fields.
* Changes actress to actor.
* changes PK values from text (nm... and tt...) to integers as they will
  consume less storage in the database and are often faster; will also
  speed up other sampling tasks for other data files

nconst (nm...) is PK
knownForTitles (tt...) is FK

```{r}
start_time <- Sys.time()

fn <- "name.basics.tsv.gz"
new.fn <- "name.basics.sample.tsv"
new.fn.kf <- "name.basics.kf.sample.tsv"

maxLines <- -1
sampleSize <- 5000

start_time_read <- Sys.time()

df.names.all <- read.csv(file = paste(path,fn,sep="/"), 
                         sep = '\t', 
                         header = TRUE, 
                         stringsAsFactors = FALSE,
                         nrows = maxLines)

n.all <- nrow(df.names.all)

end_time_read <- Sys.time()
print(paste0("Time to read file: '",
             round(difftime(end_time_read,
                            start_time_read,
                            units = "secs"),2), " sec",
             ' | ', n.all, ' lines read'))


# analysis of file

# only retain names for which there's a birthyear
df.names.all.bY <- df.names.all[which(df.names.all$birthYear != "\\N"),]

# only look at those names who contain any actor/actress role
# add column isActor to TRUE
a <- grepl( 'actor', df.names.all.bY[,5], fixed = TRUE) | 
  grepl( 'actress', df.names.all.bY[,5], fixed = TRUE)
df.names.all.bY$isActor[a] <- TRUE
df.names.actors <- df.names.all.bY[a,]


# sample actor data
set.seed(55776)
s <- round(runif(sampleSize, min = 1, max = (nrow(df.names.actors))),0)
df.names.sample <- df.names.actors[s,]

# clean data

# which() is much faster than a for loop

# eliminate NA and \\N values from fields
nas <- which(is.na(df.names.sample[,3]) | df.names.sample[,3] == "\\N")
df.names.sample[nas,3] <- 0
df.names.sample[,3] <- as.numeric(df.names.sample[,3])

nas <- which(is.na(df.names.sample[,4]) | df.names.sample[,4] == "\\N")
df.names.sample[nas,4] <- 0
df.names.sample[,4] <- as.numeric(df.names.sample[,4])

nas <- which(is.na(df.names.sample[,5]) | df.names.sample[,5] == "\\N")
df.names.sample[nas,5] <- ''

nas <- which(is.na(df.names.sample[,6]) | df.names.sample[,6] == "\\N")
df.names.sample[nas,6] <- ''

# add a column for PK that is numeric to reduce storage in a table
df.names.sample$nconstPK <- as.numeric(substr(df.names.sample[,1], 3, 99))

# save each nm plus one tt per line; the original tt's for knownFor remain
# in the file as a comma-separated field

n.sample <- nrow(df.names.sample)
print (paste("Total number of retained names: ", n.sample))

# clear memory and then garbage collect except for df.names.sample
rm(df.names.actors, df.names.all, df.names.all.bY, df.names.raw)
gc()

# splot all tts in knownFor column into list of vectors of chr
titles <- strsplit(df.names.sample[,6], ",")

# garbage collect memory
gc(verbose = F)

start_time_loop <- Sys.time()

# vectors are much faster than data frames; pre-allocate memory
v.nm <- vector(mode = "character", length = n.sample)
v.tt <- vector(mode = "character", length = n.sample)

k <- 1

for (i in 1:n.sample)
{
  if ((i %% 10000) == 0) {
    print(paste("Iteration:",i,"started..."))
  }
  
  # write each nm and tt pair to a row
  tt <- titles[[i]]
  l <- length(tt)
  if (l > 0)
  {
    for (j in 1:l)
    {
      # adding/writing to vector is much faster than dataframe
      v.nm[k] <- df.names.sample[i,1]
      v.tt[k] <- tt[j]
      
      k <- k + 1
    }
  } 
}

end_time_loop <- Sys.time()
totTm <- difftime(end_time_loop, start_time_loop, units = "secs")
print(paste0('Time to process ', s,' samples: ', 
             round(totTm, 2), " sec"))



# write sampled data to new TSV
write.table(df.names.sample, file = paste(path,new.fn,sep="/"), 
          col.names = TRUE, sep = '\t', row.names = F)

# write the PK/FK file with PKs and FKs as numbers
df.nm.tt <- data.frame(as.numeric(substr(v.nm, 3, 99)),
                       as.numeric(substr(v.tt, 3, 99)))

write.table(df.nm.tt, file = paste(path,new.fn.kf,sep="/"), 
          sep = '\t', row.names = F,
          col.names = c("nconst","tconst"))

end_time <- Sys.time()
print(paste0('Time: ', round(difftime(end_time, start_time, units = "secs"),2), " sec"))

```

## Read sampled name.basics.sample file

```{r}
start_time <- Sys.time()

new.fn <- "name.basics.sample.tsv" 

df.names <- read.csv(file = paste(path,new.fn,sep="/"), 
                     sep = '\t', 
                     header = TRUE, 
                     stringsAsFactors = FALSE)

end_time <- Sys.time()
print(paste0('Time: ', round(difftime(end_time, start_time, units = "secs"),2), " sec"))

# ensure that data was read
head(df.names)
tail(df.names)

```